### 2022-01-07 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_neutral_landscapes_jl@ad021b39d7baa9690157c03e794ab7575ed9aba8 🚀

Neutral landscapes are increasingly used in ecological and evolutionary
studies to provide a null expectation of the variance of a given metric
over space.

Wide range of disciplines: from landscape genetics {[}{]}, to spatial
ecology {[}{]}, and biogeography {[}{]}.

As biodiversity science becomes increasingly concerned with temporal
change and its consequences, its clear there is a gap generating neutral
landscapes that change over time. In this ms we present how
\texttt{NeutralLandscapes.jl} is orders of magnitudes faster than
packages \texttt{nlmpy} (in python) or \texttt{NLMR} (in R). In addition
we then present a novel method for generating landscape change with
prescribed levels of spatial and temporal autocorrelation.

@@

fig.~\ref{fig:allmethods} shows a replica of Figure 1 from
(\textbf{nlmpycite?}), which shows the capacity of the library to
generate different types of neutral landscapes, and then apply masks and
categorical classifcation to them.

Table of methods.

\begin{longtable}[]{@{}llllll@{}}
\toprule
\begin{minipage}[b]{0.35\columnwidth}\raggedright
Model\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright
nlmpy?\strut
\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedright
NLMR\strut
\end{minipage} & \begin{minipage}[b]{0.33\columnwidth}\raggedright
Description\strut
\end{minipage} & \begin{minipage}[b]{0.02\columnwidth}\raggedright
Reference\strut
\end{minipage} & \begin{minipage}[b]{0.02\columnwidth}\raggedright
Aliases\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.35\columnwidth}\raggedright
No gradient\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Each cell is drawn randomly\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Planar gradient\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
A gradient from low to high at a given angle\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Distance gradient\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Each cell is the distance between that cell and a location\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Random rectuangular cluster\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Covers the plane in random rectanges until covered\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Random element nearest-neighbor\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x*\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Discrete categories based on distance a set of \texttt{n} random
points\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\texttt{nlm\_mosaictess}, k-means\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Random cluster nearest-neighbor\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Starts with \(n\) seed points and grows clusters probabilistically\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Random curds\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Gaussian Field\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Diamond-square\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Improvement on Diamond-square and fractal brownian motion.\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Perlin noise\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
Method for noise with ``smoother'' features than DS/MPD\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.35\columnwidth}\raggedright
Mosaic random field\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.02\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

@@

\hypertarget{what-methods-have-been-called-different-things-but-are-actually-the-same-thing}{%
\subsection{What methods have been called different things but are
actually the same
thing?}\label{what-methods-have-been-called-different-things-but-are-actually-the-same-thing}}

@@

>>>>

Neutral landscapes are increasingly used in ecological and evolutionary
studies to provide a null expectation spatial variation of a given
measurement. Originally developed to simulate the spatially
autocorrelated data (\textbf{Gardner1987NeuMod?};
\textbf{Milne1992SpaAgg?}), the have seen use in a wide range of
disciplines: from landscape genetics (\textbf{Storfer2007PutLan?}), to
landscape and spatial ecology (\textbf{Tinker2004HisRan?};
\textbf{Remmel2013CatCla?}), and biogeography
(\textbf{Albert2017BarDis?}).

The two primary packages used to simulate neutral landscapes are
\texttt{NLMR} in (the \texttt{R} language) (\textbf{Sciaini2018NlmLan?})
and \texttt{NLMpy} (in Python; \textbf{Etherington2015NlmPyt?}). We
present \texttt{NeutralLandscapes.jl}, a package in \texttt{Julia} for
neutral landscapes which is faster than both above package. Here we
demonstrate that \texttt{NeutralLandscapes.jl}, depending on the method,
is orders of magnitude faster than previous neutral landscape packages.
As biodiversity science becomes increasingly concerned with temporal
change and its consequences, its clear there is a gap in methodology in
generating neutral landscapes that change over time. In addition we
present a novel method for generating landscape change with prescribed
levels of spatial and temporal autocorrelation, which is implemented in
\texttt{NeutralLandscapes.jl}

@@

fig.~\ref{fig:allmethods} shows a replica of Figure 1 from
(\textbf{Etherington2015NlmPyt?}), which shows the capacity of the
library to generate different types of neutral landscapes, and then
apply masks and categorical classifcation to them.

@@

@@

\(f_{ST}(M_{ij}) = r + v \cdot [Z(\delta)]_{ij}\)
